Linear Algebra as a Unifying Theory\\of Wireless Workload Modeling

### 2008-02-03 paper.tex
I'm not sure that the flow pairing really belongs here.
The problem is that there's no way to make sense of summing flow pairings.

The traffic workload pattern in a network is simply a collection of packets between hosts in the network sent at certain times. It is standard in network analysis, however, to aggregate sequences of packets sharing the same \caps{IP} protocol, source and destination nodes, and source and destination \caps{TCP} or \caps{UDP} port numbers. Such a sequence of packets is called a \textit{flow}. The traffic pattern is characterized completely by the detailed behavior of it's flows. The behavior of a flow is defined by seven properties: its source and destination nodes; its start and end times; the specific sizes of and intervals between its constituent packets; and finally, whether it is paired with another flow in a request-response relationship.\footnote{Flow pairing is typically not addressed in traffic models. However, the most common transport protocol, {\scriptsize{TCP}}, is full-duplex, exhibiting tightly coupled pairs of client-server flows. Therefore, accurate representation of traffic behavior requires knowing which flows are request-response pairs. Moreover, although they are not strictly full-duplex, it is generally appropriate to view {\scriptsize{UDP}} and {\scriptsize{ICMP}} flows as paired, since the client-server paradigm is still used. Each {\scriptsize{DNS}} response, for example, is naturally associated with its corresponding request, and will have symmetrical {\scriptsize{IP}} addresses and port numbers.} We propose to represent the behavior of each flow in terms of seven vectors---one for each aspect of its behavior. The aggregate behavior of all the flows in a workload pattern are then represented as matrices, the rows of which are the behavior vectors of the flows.

@@

The {packets-bytes-duration} representation is clearly overly simplistic and very incomplete. (Yet, it is more detailed than the most commonly used experimental traffic models!) It captures nothing about variability of packet sizes, nor about the ``burstiness'' of packet transmissions; it leaves start time unspecified; it says nothing about the source or destination of the flow. At the opposite extreme, it is possible to embed the exact start time of each flow, and the exact size and transmission times of its packets in a vector space, etc. The difficulty comes in doing so in a manner that makes the summation of vectors correspond naturally and meaningfully to the aggregate behavior of flows. We take an intermediate approach, suggested by the results of~\cite{Karpinski07:realism}: to reproduce network performance accurately, it suffices to know, for each flow, the correct source node, destination node, start time, end time, empirical distributions of packet sizes and inter-packet intervals for each flow, and which other flow (if any) it is a response to. If there are $f$ flows in the entire workload pattern, then its collective behavior can be described with seven matrices; the rows of each matrix are the behavior vectors of the flows, indexed in some chosen order. We use the following notation for the behavior matrices and their constituent vectors and elements:\vspace{-0.25em}
\begin{tabbing}
\hspace{1.5em}\=\bullet\hspace{0.5em}\=\hspace{9.75em}\=
$D\:$\=$=\setx{\vec{d}_i}\:$\=$=\setx{d_{ij}}\:$\=\kill
\>\bullet\> sources: \>$S$\>$=\setx{\vec{s}_i}$\>$=\setx{s_{ij}}$\>$\in\R^{f \times n}$,\\
\>\bullet\> destinations: \>$D$\>$=\setx{\vec{d}_i}$\>$=\setx{d_{ij}}$\>$\in\R^{f \times n}$,\\
\>\bullet\> flow pairs: \>$P$\>$=\setx{\vec{p}_i}$\>$=\setx{p_{ij}}$\>$\in\R^{f \times f}$,\\
\>\bullet\> start times: \>$B$\>$=\setx{\vec{b}_i}$\>$=\setx{b_{ij}}$\>$\in\R^{f \times d_t}$,\\
\>\bullet\> end times: \>$E$\>$=\setx{\vec{e}_i}$\>$=\setx{e_{ij}}$\>$\in\R^{f \times d_t}$,\\
\>\bullet\> packet sizes: \>$Z$\>$=\setx{\vec{z}_i}$\>$=\setx{z_{ij}}$\>$\in\R^{f \times d_z}$,\\
\>\bullet\> inter-packet intervals: \>$V$\>$=\setx{\vec{v}_i}$\>$=\setx{v_{ij}}$\>$\in\R^{f \times d_v}$.
\end{tabbing}

@@

Let $n$ be the number of nodes in the given traffic collection. We map nodes onto the standard basis vectors of $\R^n$: if the $j$th node is the source of the the $i$th flow, then $\vec{s}_i=\mathbf{e}_j$ (the vector with 1 at index $j$ and zeros elsewhere); if the $k$th node is the destination, then $\vec{d}_i=\mathbf{e}_k$. This may seem like an overly expansive representation, given that start and end nodes can be represented as a unidimensional indices from $\set{1,\dots,n}$. The advantage becomes clear when we consider vector operations. The sum of the source indices over a collection is meaningless; the sum of the the source vectors, however, is an $n$-dimensional histogram indicating how often each source node occurs in the collection.

\subsection{Flow Pairing: Requests \& Responses}

>>>>

%The traffic workload pattern in a network is the collection of actions taken by hosts, on behalf of users and applications, that affect the behavior of the network itself. By design, Internet Protocol (\caps{IP}) networks treat all data in the same manner~\cite{Clark88}. Therefore, the actual information transmitted is irrelevant, only the manner of transmission matters. Although networks are packet-switched, packets are logically organized into sequences of packets sharing the same \caps{IP} protocol, source and destination \caps{IP} address, and source and destination  \caps{TCP} or \caps{UDP} port numbers. Such a sequence of packets is called a \textit{flow}. Although there are a vast multitude different application types which use \caps{IP}, there are two fundamentally different behaviors of flows which we distinguish here: \caps{TCP}, which delivers data reliably through feedback and retransmissions; and \caps{UDP}, which delivers data only on a best-effort basis. While it is possible to implement reliable transmission similar to \caps{TCP} over \caps{UDP}, it is uncommon, so we make the generalization that network events, such as packet drops, do not significantly feedback to the behavior of upper layers. With \caps{TCP}, on the other hand, network events are known to affect the transport layer

The traffic workload pattern in a network is simply a collection of \caps{IP} packets between hosts in the network sent at certain times. It is standard in network analysis, however, to aggregate sequences of packets sharing the same \caps{IP} protocol, source and destination nodes, and source and destination \caps{TCP} or \caps{UDP} port numbers. Such a sequence of packets is called a \textit{flow}. The traffic pattern is characterized completely by the detailed behavior of it's flows. The behavior of a flow is defined by seven properties: its source and destination nodes; its start and end times; the specific sizes of and intervals between its constituent packets; and finally, whether it is paired with another flow in a request-response relationship.\footnote{Flow pairing is typically not addressed in traffic models. However, the most common transport protocol, {\scriptsize{TCP}}, is full-duplex, exhibiting tightly coupled pairs of client-server flows. Therefore, accurate representation of traffic behavior requires knowing which flows are request-response pairs. Moreover, although they are not strictly full-duplex, it is generally appropriate to view {\scriptsize{UDP}} and {\scriptsize{ICMP}} flows as paired, since the client-server paradigm is still used. Each {\scriptsize{DNS}} response, for example, is naturally associated with its corresponding request, and will have symmetrical {\scriptsize{IP}} addresses and port numbers.} We propose to represent the behavior of each flow in terms of seven vectors---one for each aspect of its behavior. The aggregate behavior of all the flows in a workload pattern are then represented as matrices, the rows of which are the behavior vectors of the flows.

@@

The {packets-bytes-duration} representation is clearly overly simplistic and very incomplete. (Yet, it is more detailed than the most commonly used experimental traffic models!) It captures nothing about variability of packet sizes, nor about the ``burstiness'' of packet transmissions; it leaves start time unspecified; it says nothing about the source or destination of the flow. At the opposite extreme, it is possible to embed the exact start time of each flow, and the exact size and transmission times of its packets in a vector space, etc. The difficulty comes in doing so in a manner that makes the summation of vectors correspond naturally and meaningfully to the aggregate behavior of flows. We take an intermediate approach, suggested by the results of~\cite{Karpinski07:realism}: to reproduce network performance accurately, it suffices to know, for each flow, the correct source node, destination node, start time, end time, empirical distributions of packet sizes and inter-packet intervals for each flow, and which other flow (if any) it is a response to. If there are $f$ flows in the entire workload pattern, then its collective behavior can be described with seven matrices; the rows of each matrix are the behavior vectors of the flows, indexed in some chosen order. We use the following notation for the behavior matrices and their constituent vectors and elements:\vspace{-0.25em}
\begin{tabbing}
\hspace{1.5em}\=\bullet\hspace{0.5em}\=\hspace{9.75em}\=
$D\:$\=$=\setx{\vec{d}_i}\:$\=$=\setx{d_{ij}}\:$\=\kill
\>\bullet\> sources: \>$S$\>$=\setx{\vec{s}_i}$\>$=\setx{s_{ij}}$\>$\in\R^{f \times n}$,\\
\>\bullet\> destinations: \>$D$\>$=\setx{\vec{d}_i}$\>$=\setx{d_{ij}}$\>$\in\R^{f \times n}$,\\
\>\bullet\> response: \>$R$\>$=\setx{\vec{r}_i}$\>$=\setx{r_{ij}}$\>$\in\R^{f \times f}$,\\
\>\bullet\> start times: \>$B$\>$=\setx{\vec{b}_i}$\>$=\setx{b_{ij}}$\>$\in\R^{f \times d_t}$,\\
\>\bullet\> end times: \>$E$\>$=\setx{\vec{e}_i}$\>$=\setx{e_{ij}}$\>$\in\R^{f \times d_t}$,\\
\>\bullet\> packet sizes: \>$Z$\>$=\setx{\vec{z}_i}$\>$=\setx{z_{ij}}$\>$\in\R^{f \times d_z}$,\\
\>\bullet\> inter-packet intervals: \>$V$\>$=\setx{\vec{v}_i}$\>$=\setx{v_{ij}}$\>$\in\R^{f \times d_v}$.
\end{tabbing}

@@

Let $n$ be the number of nodes in the given traffic collection. We let $s_{ij}=1$ if the $j$th node is the source of the $i$th flow, and zero otherwise; similarly, $d_{ik}=1$ if the destination of the $i$th flow is the $k$th node. This sparse matrix may seem like an overly expansive representation, given that start and end nodes can be represented as unidimensional indices from $\set{1,\dots,n}$. The advantage becomes clear when we consider vector operations combining collections of behavior vectors. The sum of the source indices over a collection is meaningless; the sum of the the source vectors, however, is an $n$-dimensional histogram indicating how often each source node occurs in the collection.

\subsection{Requests \& Responses: Flow Pairing}
